---
title: "p8105_hw1_sc5154"
author: "Shaohan Chen"
date: "2022-09-21"
output: github_document
---

This document is my solution to Homework1 of P8105 Data Science.

Load necessary packages.
```{r}
library(tidyverse)
```

## Problem 1

Install the package `palmerpenguins` containing the `penguins` dataset. (This code is not  shown or executed in RMarkdown file)

Load the `penguins` dataset.
```{r cars}
data("penguins", package = "palmerpenguins")
```

Then I'm going to write a short description of the `penguins` dataset.

Let's first take a look at part of data in dataset.
```{r}
head(penguins, 5)
tail(penguins, 5)
```

From the data displayed above, we can see that `penguins` dataset is describing features of a group of penguins. It has 8 columns, each represents a feature variable of penguins. The variables include `species`, `flipper_length_mm`,`sex` and `year` etc. 

Here's a detailed summary of statistics for all data in the dataset.
```{r}
skimr::skim(penguins)
```

Or a brief version of summary towards the 8 variables in dataset:
```{r}
summary(penguins)
```

Next I'll display the size of the dataset.

The row number of the dataset is:
```{r}
nrow(penguins)
```

The column number of the dataset is:
```{r}
ncol(penguins)

```

Then I'll calculate the mean flipper_length:
```{r}
mean(penguins$flipper_length_mm)
```

The result of mean is "NA", which indicates there may be missing values in the dataset. 

But we can still calculate the mean of the rest values ignoring the values that are not available:
```{r}
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

Last I'll make a scatterplot of `flipper_length_mm(y)` vs `bill_length_mm(x)`, and color points using the `species` variable.
```{r}
plot_df = tibble(
  x = penguins$bill_length_mm,
  y = penguins$flipper_length_mm
)

scatterplot1 = ggplot(plot_df, aes(x = x, y = y, color = penguins$species)) + geom_point()
print(scatterplot1)
```

And I save this scatterplot to my project directly using `ggsave`.
```{r}
ggsave("scatterplot1.png", scatterplot1)
```

## Problem 2

Let's create a data frame comprised of the required variable types.
```{r}
data_vt = tibble(
  x_1 = rnorm(10, mean = 0, sd = 1),
  x_2 = x_1 > 0,
  x_3 = c('a','b','c','d','e','f','g','h','i','j'),
  x_4 = factor(c('a','a','a','a','b','b','b','b','c','c'),
        levels = c('a','b','c'))
)
```

Next I'll try to take the mean of each variable in dataframe using the `pull` function.
```{r}
mean(pull(data_vt, var = 1))
mean(pull(data_vt, var = 2))
mean(pull(data_vt, var = 3))
mean(pull(data_vt, var = 4))
```

As shown on the output above, the mean works on random sample and logical vector (x_1 and x_2), but not on character vector or factor vector(x_3 and x_4).

Finally, I'll apply the `as.numeric` function to the logical, character, and factor variables (But hide the output).
```{r, results = "hide"}
as.numeric(pull(data_vt, var = 2))
as.numeric(pull(data_vt, var = 3))
as.numeric(pull(data_vt, var = 4))

```

It seems that `as.numeric` works on the logical and factor variables(x_2 and x_4), but not on character variable(x_3).

That indicates me that Rstudio consider logical variable as `1` or `0` in numeric calculation, so I can take the mean of the logical vector. 

Meanwhile, Rstudio can modify the format of logical variable into `0` and `1`, and also change factor vector as a numeric vector, but cannot change character vector into numeric vector. And when Rstudio cannot transform the format of characer vector, it just fill the vector with `NA`.

That's why `as.numeric` works on the logical and factor variables(x_2 and x_4), but not on character variable(x_3). And if we want to calculate with factor vector, we have to manually transform it into numeric vector.


That's all for Homework1. Thank you for your time!
