---
title: "p8105_hw1_sc5154"
author: "Shaohan Chen"
date: "2022-09-21"
output: github_document
---

This document is my solution to Homework1 of P8105 Data Science.

Load necessary packages.
```{r}
library(tidyverse)
```

## Problem 1

Install the package `palmerpenguins` containing the `penguins` dataset. (This code is not  shown or executed in RMarkdown file)

Load the `penguins` dataset.
```{r cars}
data("penguins", package = "palmerpenguins")
```

Then I'm going to write a short description of the `penguins` dataset.

Let's first take a look at part of data in dataset.
```{r}
head(penguins, 10)
tail(penguins, 10)
```

And the whole dataset:
```{r}
penguins
```

From the data displayed above, we can see that `penguins` dataset is describing features of a group of penguins. It has 8 columns, each represents a feature variable of penguins. The variables include `species`, `flipper_length_mm`,`sex` and `year` etc. 

Here's a detailed summary of statistics for all data in the dataset.
```{r}
skimr::skim(penguins)
```

Or a brief version of summary towards the 8 variables in dataset:
```{r}
summary(penguins)
```

Next I'll display the size of the dataset.

The row number of the dataset is:
```{r}
nrow(penguins)
```

The column number of the dataset is:
```{r}
ncol(penguins)

```

Then I'll calculate the mean flipper_length:
```{r}
mean(penguins$flipper_length_mm)
```

The result of mean is "NA", which indicates there may be missing values in the dataset. 

But we can still calculate the mean of the rest values ignoring the values that are not available:
```{r}
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

Last I'll make a scatterplot of `flipper_length_mm(y)` vs `bill_length_mm(x)`, and color points using the `species` variable.
```{r}
plot_df = tibble(
  x = penguins$bill_length_mm,
  y = penguins$flipper_length_mm
)

scatterplot1 = ggplot(plot_df, aes(x = x, y = y, color = penguins$species)) + geom_point()
print(scatterplot1)
```

And I save this scatterplot to my project directly using `ggsave`.
```{r}
ggsave("scatterplot1.png", scatterplot1)
```
